\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{dirtree}
\usepackage{listings}
\usepackage{ amssymb }
\usepackage{ textcomp }
\newcommand\tab[1][1cm]{\hspace*{#1}}
\title{Lógica proposicional\\(5ª semana)}
\begin{document}
\maketitle
\pagebreak
\graphicspath{/mnt/data_1/code/asignaturas/sint/practicas/pr4}


\section{ejercicio 8.6}
\subsection{enunciado}
epresent the following  sentences in first-order logic,  using  a  consistent  vocabulary (which you must define):
\begin{enumerate}
	\item Some students took French in spring 2001.
	\item Every student who takes French passes it.
	\item Only one student took Greek in spring 2001.
	\item The best score in Greek is always higher than the best score in French.
	\item Every person who buys a policy is smart.
	\item No person buys an expensive policy.
	\item There is an agent who sells policies only to people who are not insured.
	\item There is a barber who shaves all men in town who do not shave themselves.
	\item A person born in the UK, each of whose parents is a IJK citizen or a UK resident, is a UK citizen by birth.
	\item A person born outside the UK, one of whose parents is a UK citizen by, acitizen by descent.
	\item Politicians can fool some of the people all of time, they can fool all of people some of the time, but they can't fool all of the people all
	of the time.
\end{enumerate}
\subsection{resolución}
\begin{enumerate}
	\item $\exists$ s isStudent(s) $\rightarrow$ $\exists$ c tookClases(s,c) : where c is french
	\item $\forall$ s isStudent(s) $\land$ $\exists$ c tookClases(s,c) $\rightarrow$ $\exists$ c passClass(s,c) : where c is french
	\item $\exists$ s isStudent(s) $\land$ $\exists$ g tookClass(s, g) $\rightarrow$ $\exists!$ study(s, g) : where s is student and g is Greek
	\item $\forall$ c isCalificationOfGreek(g) $\land$ $\forall$ f isCalificationOfFrench(f) $\rightarrow$ isHigher(g,f) : where the functions evaluates if param1 $\gg$ param2
	\item $\exists$ p isPerson(p) $\land$ $\exists$ l isPolicy(l) $\rightarrow$ $\forall$ buys(p, l) $\rightarrow$ isWhise(p)
	\item $\nexists$ p isPerson(p) $\land$ $\exists$ l isPolicy(l) $\rightarrow$ buys(p, l)
	\item $\exists$ p isPerson(p) $\land$ $\exists$ a isAgent(a) $\land$ $\exists$ l isPolicy(l) $\rightarrow$ buys(a, l) $\land$ buysTo(a,p)
	\item $\exists$ p isPerson(p) $\land$ $\exists$ b isBarber(b) $\land$ $\exists$ t isTown(t) $\rightarrow$ dontShaveHimself(p) $\land$ isSaidTown(t) $\rightarrow$ shaves(b, p)
	\item $\exists$ p1 isPerson(p1) $\land$ $\exists$ p1 isPerson(p1) $\land$ $\exists$ s isPerson(s) $\rightarrow$ bornInUK(p1) $\land$ bornInIJK(p2) $\land$ bornInUK(s) $\rightarrow$ haveUKNationality(s)
	\item $\exists$ p1 isPerson(p1) $\land$ $\exists$ p1 isPerson(p1) $\land$ $\exists$ s isPerson(s) $\rightarrow$ bornInUK(p1) $\land$ bornInUK(p2) $\land$ notBornInUK(p1) $\rightarrow$ haveUKNationality(s)
	\item $\exists$ Po isPolitician(po) $\land$ $\exists$ Pe isPerson(pe) $\land$ $\forall$ All isPerson(all) $\rightarrow$ ( foolAllOfTheTime(po, pe) $\lor$ foolSomeOfTheTime(po, all) ) $\land$ $\lnot$ foolAllOfTheTime(po, all)
\end{enumerate}

\section{ejercicio 8.7}
\subsection{enunciado}
peresent the sentence " All Germans speak the same languages " in predicate calculus. Use Speaks (x, 1) , meaning that person x speaks language 1.
\subsection{resolución}
$\exists$ x isPerson(x) $\rightarrow$ isGerman(x) $\land$ speaks(x, 1)

\section{ejercicio 8.8}
\subsection{enunciado}
What axiom is needed to infer the fact Female(Laura) is given the facts Male(Jim) and Spouse(Jim, Laura)?
\subsection{resolución}
$\exists$ Laura $\land$ $\exists$ Jim Spouse(Jim, Laura) $\rightarrow$ isFemale(Laura) $\land$ isMale(Jim)

\section{ejercicio 8.9}
\subsection{enunciado}
Write a  general  set of facts  and  axioms to represent ithe assertion " Wellington heard about  Napoleon's  death " and  to  correctly  answer  the ques.tion " Did Napoleon  hear  about Wellington's death? "
\subsection{resolución}
$\exists$ Napoleon $\land$ $\exists$ Wellingtong $\rightarrow$ heardAbowtTheDeathOf(Wellington, Napoleon)
\\\tab answer: yes

\pagebreak

\section{ejercicio 8.10}
\subsection{enunciado}
Rewrite the propositional wumpus world facts from Section 7.5 into first - order logic. How much more compact is this version?
\subsection{resolución}
No he encontrado a que hace referencia

\section{ejercicio 8.10}
\subsection{enunciado}
Write axioms  describing  the  predicates Grandchild, GreatGrandparent, Brother, Sister, Daughter, Son,  Aunt, Uncle, BrotherInLaw, SisterInLaw, and Firstcousin. Find  out  the  proper  definition of nzth cousin n times  removed,  and  write  the  definition in first - order logic. Now write down the basic facts depicted in the family tree in Figure 8.5. Using a suit - able logical  reasoning  system, T E L L it all the  sentences you have written down,  and A S K it who are Elizabeth's  grandchildren, Diana's  brothers - in - law, and Zara's  great - grandparents.
\subsection{resolución}
( $\exists$ (George, Mum)
$\land$ $\exists$ (Spencer, Kydd, Elizabeth, Philip, Margatet)
$\land$ $\exists$ (Diana, Charles, Anne, Mark, Andrew, Sharah, Edward)
$\land$ $\exists$ (William, Harry, Peter, Zara, Beatrice, Eugenie) )\\
$\land$ $\forall$ (x, y, y') sonOf(x, y, y') $\rightarrow$ ( isPerson(x) $\land$ isPerson(y) $\land$ isPerson(y') $\land$ isParent(y) $\land$ isParent(y')) $\land$ isSon(x))\\
$\rightarrow$\\
( isSon(Elizabeth, george, Mun) $\land$ isSon(Margatet, george, Mun))\\
$\land$
( isSon(Diana, Spencer, Kydd) $\land$ isSon(Charles, Elizabeth, Philip)\\ $\land$ isSon(Andrew, Elizabeth, Philip) $\land$ isSon(Edward, Elizabeth, Philip) )\\
$\land$ (isSon(William, Diana, Charles) $\land$ isSon(Harry, Diana, Charles) $\land$ isSon(Peter, Annie, Mark) $\land$ isSon(Zara, Annie, Mark) $\land$ isSon(Beatrice, Beatrice, Eugenie) $\land$ isSon(Eugenie, Beatrice, Eugenie) )

\section{ejercicio 8.11}
\subsection{enunciado}
Write down a sentence asserting that + is a commutative function. Does your sentence follow from the Peano axioms? If so, explain why; if not, give a model in which the axioms are true and your sentence is false.
\subsection{resolución}


\pagebreak

\section{ejercicio 1}
Obtén la forma normal clausulada de las siguientes fórmulas:
\subsection{enunciado}
\begin{enumerate}
	\item (p $\land$ q) $\lor$ (r $\land$ s)
	\item (p $\land$ q) $\rightarrow$ r
	\item (p $\land$ q) $\oplus$ r
	\item ($\neg$(p $\land$ $\neg$q) $\lor$ r] $\rightarrow$ [(p  $\land$ $\neg$r) $\lor$ q]
	\item (p $\land$ q) $\rightarrow$ r] $\iff$ [(p $\land$ $\neg$r)  $\rightarrow$ $\neg$ q]

\end{enumerate}
\subsection{resolución}
\begin{enumerate}
	\item (p $\land$ q) $\lor$ (r $\land$ s)
	\item $\neg$(p $\land$ q) $\lor$ r
	\item ((p $\land$ q) $\land$ r) $\lor$ $\neg$(p $\land$ q) $\land$ $\neg$r
	\item $\neg$[$\neg$(p $\land$ $\neg$q) $\lor$ r] $\lor$ [(p  $\land$ $\neg$r) $\lor$ q]
	\item .[$\neg$[$\neg$(p $\land$ q) $\lor$ r] $\lor$ [$\neg$(p $\land$ $\neg$r)  $\lor$ $\neg$ q] ] $\land$ [ $\neg$ [$\neg$(p $\land$ $\neg$r)  $\lor$ $\neg$ q] $\lor$ [$\neg$(p $\land$ q) $\lor$ r] ]
\end{enumerate}

\section{ejercicio}
\subsection{enunciado}
Representa los siguiente hechos con lógica de predicados:
\begin{enumerate}
	\item Algunas plantas no tienen flores
	\item Cualquier edificio es habitable
	\item No hay delito sin causa
	\item Algunas personas son insoportables
	\item Existen personas que no comen carne
	\item No es oro todo lo que reluce
	\item Ningún asesino es bondadoso
	\item El que estudia, aprueba
	\item No todos los animales son racionales
	\item Existen personas que aman a todo el mundo
	\item No es verdad que todas las personas no amen a todo el mundo
\end{enumerate}
\subsection{resolución}
\begin{enumerate}
	\item $\exists$ Plant esPlanta(Plant) $\rightarrow$ haveNoFlower(Plant)
	\item $\forall$ Edificio esEdificio(Edificio) $\rightarrow$ esHabitable(Edificio)
	\item $\forall$ Delito esDelito(Delito) $\rightarrow$ conCausa(Delito)
	\item $\exists$ Persona esPersona(Persona) $\rightarrow$ esInsoportable(Persona)
	\item $\exists$ Persona esPersona(Persona) $\rightarrow$ $\neg$ comeCarne(Persona)
	\item $\exists$ objeto esBrillante(objeto) $\rightarrow$ $\neg$ esOro(objeto)
	\item $\forall$ Asesino Es persona(Asesino) $\land$ Asesino(Asesino) $\rightarrow$ $\neg$ esBondadoso(Asesino)
	\item $\forall$ Persona esPerosna() $\land$ estudia(persona) $\rightarrow$ aprueba(Persona)
	\item $\exists$ Animal esAnimal(Animal) $\rightarrow$ $\neg$ esRacional(Animal)
	\item $\exists$ Persona sPersona(Perosna) $\land$ $\forall$ Mundo esPersona(Mundo) $\rightarrow$ amar(Persona, Mundo)
	\item $\neg$$\forall$ Persona siPersona(Persona) $\land$ $\forall$ Mundo esPersona(Mundo) $\rightarrow$ $\neg$ amar(Persona, Mundo)
\end{enumerate}


\vspace*{\fill}
\raggedleft Documento escrito en \LaTeX{}
\end{document}
